\pdfbookmark[1]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer

%\begin{flushright}{\slshape
%Alles ist relativ} \\ \medskip
%\end{flushright}
%\bigskip
\begingroup

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}
Before becoming all technical, I would like to express my sincere gratitude to the many people who were involved in the writing of this thesis and helped me get to the finish line of my master's studies.

I was given a lot of freedom to choose the course of my research and my topic, and I am thrilled that I could pursue it this way:
A bit more than a year ago,  I was seriously contemplating about quitting my studies at HPI for a university that would have allowed me to learn more about machine learning.
Thanks to this thesis, I could not only remain at HPI, but also pursue my learning goal, and learn a lot about the structure and wording of a scientific paper along the way. My supervisors Luise Pufahl and Haojin Yang were always there to assist me with literature work and technical advice.

A range of people helped me in small but essential ways:
Christian Bartz helped me choose a suitable deep-learning framework.
The authors Stefan Schönig and Richard Jasinski provided help in reverse-engineering the comparison models and were also helpful in offering advice.
Willi Gierke helped me with some fundamental questions on LSTM.
Bernhard Rabe and Tobias Papke kept the FutureSOC running while I was doing my experiments.

Furthermore, I am happy to have had several pairs of watchful eyes proofread this document:
Erick Godínez, Thomas Goerttler, Marvin Bornstein, Stephan Detje, Georg Berecz, and Luise Pufahl.\\

Finally, I would like to express my deep gratitude to the following parties:

Stephan Lessmann, for his \textit{Business Analytics and Data Science} lecture, that sparked in me the interest to pursue this topic. Without it, I would not have found an interest in analytics.

Hasso Plattner, for founding the institute that has made so great of an impact on my life, it could not be summed up on a single page.
I am very thankful to him, and I hope that I get the chance to express my gratitude personally one day.

Important people stood in the background since my start at HPI in 2012 and provided advice, infrastructure, warmth, and every student's favorite: free money. Dear Caroline, Stefan, and Adrian Wolff, dear Gerda Mikus - thank you for sharing life with me the way you do.
\endgroup
